---
title: "20241107_linear_models"
output: github_document
date: "2024-11-07"
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

```

In airbnb dataset, see how different rental prices differ

Doing some data cleaning first

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>%  
  mutate(stars = review_scores_location / 2) %>%  
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>%  
  filter(borough != "Staten Island") %>%  
  select(price, stars, borough, neighborhood, room_type) %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

## Fit some models 

consider price as an outcome that may depend on rating We fit that initial model in the following code.

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

```

Giving us:
Coefficients:
(Intercept)        stars  
     -66.50        44.11  
     
other things that we can ask 

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

broom::tidy(fit) #tidys up the summary into a tibble, easily interpret. since tibble can tidy it 

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)

fit %>% 
  broom::glance()
```

Try a bit more complex 

```{r}
fit = 
  lm(price ~ stars + borough, data = nyc_airbnb)
```

borough broken up into one reference category (the bronx) and three remaining boroughs are compared to the bronx (ordered based alphabetical)

```{r}
fit = 
  lm(price ~ stars + borough, data = nyc_airbnb)

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) %>% 
  knitr::kable(digits = 3)

```

But then change with new order of the nyc_airbnb data we factored/ reordering the borough based on frequency, so Manhatten became reference 

But still can compare Brooklyn and Queens by just comparing the two intercepts. Nothign inherently changes, just the reference 

## Some Diagnostics

```{r}
nyc_airbnb %>% 
  ggplot(aes(x = stars, y = price)) + 
  geom_point() +
  stat_smooth(method = "lm")
```

Does the stars go up as the price goes up?
Looking at this plot this is about right (but there are extreme outliers; residuals)


Regression diagnostics can identify issues in model fit, especially related to certain failures in model assumptions. Examining residuals and fitted values are therefore an imporant component of any modeling exercise.

Most diagnostics use residuals 

```{r}
modelr::add_residuals(nyc_airbnb, fit)
```

Now have a dataframe and can do dataframe things to it 


```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = resid)) +
  geom_histogram()
```


```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% 
  ggplot(aes(x = borough, y = resid)) +
  geom_violin() + 
  ylim(-200, 200)
```

See that residuals are relatively skewed and the residual distribution different across different boroughs 

This gives us an understanding of how our model regression is fitting our data. Whether there is bias, fitting one borough better than another etc. 

Residuals against star ratings 

```{r}
modelr::add_residuals(nyc_airbnb, fit) %>% # in this line of code, use modelr fit this dataframe and use fit to add residuals
  ggplot(aes(x = stars, y = resid)) +
  geom_point() 
```


Can compare residuals against fitted values

```{r}
nyc_airbnb %>% 
  modelr::add_residuals(fit) %>% 
  modelr::add_predictions(fit) %>% 
  ggplot(aes(x = pred, y = resid)) + 
  geom_point()
```

## Hypothesis Testing 

for single coefficients, looks at the table

```{r}
fit %>% 
  broom::tidy()

fit_null = lm(price ~ stars + borough, data = nyc_airbnb)
fit_alt = lm(price ~ stars + borough + room_type, data = nyc_airbnb)
```


Looking at multiple use ANOVA comparisons, when adding more than one coefficient, use ANOVA
Look at both 

```{r}
fit_null %>% 
  broom::tidy()

fit_alt %>% 
  broom::tidy()
```

ANVOA: Only works for nested models where alternative is only a small deviation from null (that null only differs from alternative by not having one thing which here is room_type)

```{r}
anova(fit_null, fit_alt) |> 
  broom::tidy()
```


## Do effects differ across boroughs?

Does the effect of increasing stars in Manhatten change the price in one way, but does it do something else in Brooklyn etc. 

First, use a lot of interactions 

```{r}
nyc_airbnb %>% 
  lm(price ~ stars * borough + room_type * borough, data = .) %>% 
  broom::tidy() 
```

see that p-values are quite significant so some of these interactions matter 

But other times not looking specifically at statisitcs but just want to fit a general model, one in The Bronx, another in Brooklyn, etc

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy() 

nyc_airbnb %>% 
  filter(borough == "Brooklyn") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy() 
```

linear model of price against stars and roomtype
i just want to know the effect of room and stars in Manhattan only 
27.1 effect of increasing starts on price for Manhattan 
$27 increase price for stars

then for Brookyln it is $21
and The Bronx for $4.45

Getting fancy and using list cols

```{r}
nyc_airbnb %>% 
  nest(data = -borough) #now getting different dataset for each borough, then can do the modelling for each seperately
```

```{r}
nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) %>% 
  select(borough, results) %>% 
  unnest(results)
```

mapping just means to take this column (model) and apply (broom::tidy) to everything that column 

To make the table easier to read (make it un-tidy) %>% 
```{r}
nyc_airbnb %>% 
  nest(data = -borough) %>% 
  mutate(
    model = map(data, \(x) lm(price ~ stars + room_type, data = x)),
    results = map(model, broom::tidy)
  ) %>% 
  select(borough, results) %>% 
  unnest(results) %>% 
  select(borough, term, estimate) %>% 
  pivot_wider(
    names_from = term,
    values_from = estimate
  )
```

See that going from whole apartment to private room in The Bronx saves $52, in Manhattan it will save $124
Going from private to shared room in Manhattan will save $30 

What about room type across Manhattan neighborhoods?

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan", 
         neighborhood == "Chinatown") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()
```

Decrease price from private room to shared room in chinatown (save $34)

```{r}
nyc_airbnb %>% 
  filter(borough == "Manhattan", 
         neighborhood == "Chelsea") %>% 
  lm(price ~ stars + room_type, data = .) %>% 
  broom::tidy()
```

Decrease price from private room to shared room in chinatown (save $20)

But this is doing it individually for each neighbourhood in Manhattan, what if i want to do it all together (using list cols again)

And the chunk below shows neighborhood-specific estimates for the coefficients related to room type.

```{r}
manhattan_airbnb =
  nyc_airbnb |> 
  filter(borough == "Manhattan")

manhattan_nest_lm_res =
  manhattan_airbnb |> 
  nest(data = -neighborhood) |> 
  mutate(
    models = map(data, \(df) lm(price ~ stars + room_type, data = df)),
    results = map(models, broom::tidy)) |> 
  select(-data, -models) |> 
  unnest(results)
```


Look at the effect of roomtype 

```{r}
manhattan_nest_lm_res %>% 
  filter(str_detect(term, "room_type")) |> 
  filter(estimate > 0)
```


```{r}
manhattan_nest_lm_res |> 
  filter(str_detect(term, "room_type")) |> 
  ggplot(aes(x = neighborhood, y = estimate)) + 
  geom_point() + 
  facet_wrap(~term) + 
  theme(axis.text.x = element_text(angle = 80, hjust = 1))
```

In general shared rooms and private rooms cost less than entire apartments
shared rooms less than private room 
(one outlier in Noho of shared rooms, but that is due to the fact there is only one shared room in Noho)


