---
title: "20241107_linear_models"
output: github_document
date: "2024-11-07"
---

```{r setup, include=FALSE}
library(tidyverse)
library(p8105.datasets)

set.seed(1)

```

In airbnb dataset, see how different rental prices differ

Doing some data cleaning first

```{r}
data("nyc_airbnb")

nyc_airbnb = 
  nyc_airbnb %>%  
  mutate(stars = review_scores_location / 2) %>%  
  rename(
    borough = neighbourhood_group,
    neighborhood = neighbourhood) %>%  
  filter(borough != "Staten Island") %>%  
  select(price, stars, borough, neighborhood, room_type) %>% 
  mutate(
    borough = fct_infreq(borough),
    room_type = fct_infreq(room_type)
  )
```

## Fit some models 

consider price as an outcome that may depend on rating We fit that initial model in the following code.

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

```

Giving us:
Coefficients:
(Intercept)        stars  
     -66.50        44.11  
     
other things that we can ask 

```{r}
fit = lm(price ~ stars, data = nyc_airbnb)

summary(fit)

names(summary(fit))

coef(fit)

broom::tidy(fit) #tidys up the summary into a tibble, easily interpret. since tibble can tidy it 

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value)

fit %>% 
  broom::glance()
```

Try a bit more complex 

```{r}
fit = 
  lm(price ~ stars + borough, data = nyc_airbnb)
```

borough broken up into one reference category (the bronx) and three remaining boroughs are compared to the bronx (ordered based alphabetical)

```{r}
fit = 
  lm(price ~ stars + borough, data = nyc_airbnb)

fit %>% 
  broom::tidy() %>% 
  select(term, estimate, p.value) %>% 
  mutate(
    term = str_replace(term, "borough", "Borough: ")
  ) %>% 
  knitr::kable(digits = 3)

```

But then change with new order of the nyc_airbnb data we factored/ reordering the borough based on frequency, so Manhatten became reference 

But still can compare Brooklyn and Queens by just comparing the two intercepts. Nothign inherently changes, just the reference 

